---
title: "App.Stats-Project"
output: word_document
---

```{r}
rm(list = ls())
# Installing packages & library
if (!require('MASS')) install.packages('MASS');
if (!require('corrplot')) install.packages('corrplot');
if (!require('leaps')) install.packages('leaps');
if (!require('car')) install.packages('car');
library(corrplot)
library(MASS)
library(leaps)
library(car)
```

```{r}
# Reading the data

linthrust_data = read.table("C:/Users/sachi/Documents/R_datafiles/LINTHALL.txt", sep = "\t", header = TRUE)
linthrust_data = linthrust_data[1:45,4:18]


# Part I
ols_model = lm(BIO~., data = linthrust_data)
summary(ols_model)

matrix=cor(linthrust_data[,2:15])
corrplot(matrix)
```

```{r}
eigen1 = eigen(matrix)
eigen_values = eigen1$values
eigen_vectors = eigen1$vectors
sum(1/eigen_values)
# As the sum of reciprocals of eigen value is 195.9633 >5*14(predictor variables) collinearity is present in data

kappas = sqrt(eigen_values[1]/eigen_values)  # condition indices
kappas[length(kappas)]    # condition number
sum(kappas>15)            # sets of collinearity
```

```{r}
# Part II
pca_model = prcomp(linthrust_data, scale. = TRUE)

var_explained = pca_model$sdev^2 / sum(pca_model$sdev^2)
plot(var_explained, xlab="PC", ylab = "Proportion of variance explained", ylim = c(0,1), type='b', main='Screen plot')
# we will only consider first 11 PC's as the proportion of variance explained from the 12th PC is very less.
s = 'Y~C1'
dfx_scaled = scale(linthrust_data[,2:15])
V = pca_model$rotation[2:15,2:15]
C = data.frame(as.matrix(dfx_scaled)%*%V)
C$Y =scale(linthrust_data$BIO)
names(C)= c('C1','C2','C3','C4','C5','C6','C7','C8','C9','C10','C11','C12','C13','C14','Y')
p =14
theta_mat= matrix(nrow=p, ncol=(2+p))
i=1
while (i<=p) {
  pcr = lm(as.formula(paste(s,'-1')), data = C)
  alpha = as.matrix(pcr$coefficients)
  theta = V[,1:i]%*%alpha
  r2 = summary(pcr)$r.squared
  theta_mat[i,1]=i
  theta_mat[i,2]=r2
  theta_mat[i,3:(p+2)]=theta
  s= paste(s,sprintf('+C%d',i+1))
  i= i+1
  }
theta_mat =as.data.frame(theta_mat)
names(theta_mat) = c('ncomp','R^2',paste('theta',1:p,sep=''))
theta_mat
# We will consider first 11 PC as after that there is barely any change/ increase in R^2.

# Calculation of Betas from PC's
pcr.full=lm(Y~.-1, data = C)
alpha=pcr.full$coefficients
theta = V%*%alpha
sy= sqrt(sum((linthrust_data$BIO-mean(linthrust_data$BIO))^2))
beta = theta* (sd(linthrust_data$BIO)/apply(linthrust_data[,2:15],2,sd))
x_bar_beats=0
for (i in 1:14) {
  b=sum(beta[i]*mean(linthrust_data[,i+1]))
  x_bar_beats = b+x_bar_beats
}

beta_0 = mean(linthrust_data$BIO) - x_bar_beats
print(beta_0)
beta
```


```{r}
# Part III
linth_data = read.table("C:/Users/sachi/Documents/R_datafiles/LINTH-5.txt", sep = "\t", header = TRUE)
linth_data = linth_data[,4:9]


#1] Stepwise Regression
full.model <- lm(BIO ~., data = linth_data)
# Model with 1 feature
model1_sal = lm(BIO~SAL, data = linth_data)
model1_pH = lm(BIO~pH, data = linth_data)
model1_k = lm(BIO~K, data = linth_data)
model1_Na = lm(BIO~Na, data = linth_data)
model1_Zn = lm(BIO~Zn, data = linth_data)
# Single variable model Summary
summary(model1_sal)
summary(model1_pH)
summary(model1_k)
summary(model1_Na)
summary(model1_Zn)

```
As the p-value of pH feature is the smallest it's the most significant variable hence it will be added as the first feature into our model
BIO ~ pH
```{r}
# Model with 2 feature 1st feature is pH 
model2_sal = lm(BIO~pH+SAL, data = linth_data)
model2_k = lm(BIO~pH+K, data = linth_data)
model2_Na = lm(BIO~pH+Na, data = linth_data)
model2_Zn = lm(BIO~pH+Zn, data = linth_data)
# 2 variable model Summary
summary(model2_sal)
summary(model2_k)
summary(model2_Na)
summary(model2_Zn)

```
As the P-value of Na feature is the smallest, the model containing ph+ Na is the most significant and thus this model will be considered
BIO~ pH + Na
```{r}
# model with 3 features BIO~ Ph+ Na+ 3RD FEATURE
model3_sal = lm(BIO~pH+Na+SAL, data = linth_data)
model3_k = lm(BIO~pH+Na+K, data = linth_data)
model3_Zn = lm(BIO~pH+Na+Zn, data = linth_data)
# 3 variable model Summary
summary(model3_sal)
summary(model3_k)
summary(model3_Zn)

```
As it can be seem from the summary by adding 3rd variable, the 3rd feature is not significant even for 0.15 significance level.
Hence the final model as only 2 features, BIO~ pH + Na
```{r}
#2]Best Subset
subset_model = regsubsets(BIO~., data = linth_data,, nbest = 10, method = "exhaustive", nvmax = 2)
with(summary(subset_model), data.frame(rsq, adjr2, cp, rss, outmat))
# By looking at the summary table Adj_R^2 and Cp values, we can narrow it down to 2 models
# BIO~ pH + Na   & BIO ~  pH + K as their Cp values are close to the the number of features+1(p+1) and they have highest Adj_R^2

# Breaking the tie with VIF.
vif(model2_Na)  # BIO~ pH + Na
vif(model2_k)   # bio~ PH + k
# As we can see, Since VIF values of features pH and Na are higher that pH+K also Cp value of model with
# pH+ K is equal to 3 that equals p+1 but on the other hand Cp value of pH and Na is closer to 2.
# Hence we can proceed with model with 2 features pH +k as there is barely any difference in the R^2 value
# BIO~ ph + k
```

